---
title: "Volume sampled plots for NES LTER bongo samples"
author: "Alexandra Cabanelas Bermudez"
date: "2024-10-03"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=4)
```

```{r echo = FALSE, include = FALSE}
#PACKAGES
library(here)
library(tidyverse)
library(tidyr)
library(RColorBrewer)
library(ggthemes)
library(ggpubr)
library(cowplot)
```

```{r echo=FALSE}
#created in bongoEventLogMerge.R
towTimes <- read.csv(file.path("output",
                               "nes-lter-zooplankton-tow-metadata-v2.csv"),
                     header = T)
```

## Calculate total time elapsed per tow
```{r include = FALSE}
# need to convert time to seconds 

# convert TIME columns to POSIXct format
towTimes$date_time_start_UTC <- as.POSIXct(towTimes$date_time_start_UTC, 
                                           format="%Y-%m-%d %H:%M:%S", 
                                           tz="UTC")
towTimes$date_time_end_UTC <- as.POSIXct(towTimes$date_time_end_UTC, 
                                         format="%Y-%m-%d %H:%M:%S", 
                                         tz="UTC")
```

```{r}
towTimes$seconds_elapsed <- ifelse(
  towTimes$date_time_end_UTC < towTimes$date_time_start_UTC, 
  as.numeric(difftime(towTimes$date_time_end_UTC + 24 * 60 * 60, 
                      towTimes$date_time_start_UTC, 
                      units = "secs")),
  as.numeric(difftime(towTimes$date_time_end_UTC, 
                      towTimes$date_time_start_UTC, 
                      units = "secs"))
)
```

### Tow length (time) Plots

```{r echo=FALSE, include=FALSE}
towTimes$station <- factor(towTimes$station, 
                           levels = c("MVCO", "L1", "L2", "L3", "L4", "L5", 
                                      "L6", "L7", "L8", "L9", "L10", "L11", "u11c"))
```

```{r echo = FALSE, fig.width=14, fig.height=10}
ggplot(towTimes %>% 
         filter(!is.na(seconds_elapsed), station != "u11c"), 
       aes(x = station, 
           y = seconds_elapsed)) +
  geom_bar(stat = "identity", 
           position = position_dodge()) +
  theme_minimal() +
  labs(y = "Elapsed Time (seconds)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) +
  facet_wrap(~ cruise, scales = "free_y", ncol = 5) 
```

```{r include = FALSE}
towTimes_long <- towTimes %>%
  pivot_longer(cols = c(seconds_elapsed), 
               names_to = "elapsed_time_type", 
               values_to = "elapsed_time")
unique_cruises <- unique(towTimes$cruise)
```


```{r warning = FALSE, echo = FALSE, fig.width=8, fig.height=4}
ggplot(subset(towTimes_long, station != "u11c"), 
       aes(x = station, y = elapsed_time, color = cruise)) +
  geom_point(size = 2.3) +
  theme_minimal() +
  labs(title = "Tow total time", y = "Elapsed Time (seconds)") +
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 11, color = "black"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15, color = "black"), 
        axis.text.x = element_text(size = 15, color = "black"), 
        axis.text.y = element_text(size = 12, color = "black"), 
        plot.title = element_text(size = 18, face = "bold", color = "black")) 

ggplot(subset(towTimes_long, station != "u11c"), 
       aes(x = station, y = elapsed_time)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Tow total time", y = "Elapsed Time (seconds)") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15, color = "black"), 
        axis.text.x = element_text(size = 15, color = "black"), 
        axis.text.y = element_text(size = 12, color = "black"), 
        plot.title = element_text(size = 18, face = "bold", color = "black")) 
```





## Calculate Volume of water sampled 

### Using flowmeter values 

Volume Sampled = (Flow End - Flow Start) * Flow calibration factor * Gear Area (m^2)

- Flowmeter calibration factor = 0.26873
- Gear Area = 0.2922 m^2
   
- tot_flow_counts_mesh <- flowmeter_end - flowmeter_start  # total counts
- revolutions <- total_counts / 10   # counts to revolutions
- total flow <- revolutions * 26873   # standard speed rotor constant

### Using ship speed & sampling time

V=(A) (T) (S)

where:

- V = calculated volume of water filtered (in meters cubed)
- A = area of net mouth (in meters squared) = 0.2922
- T = duration of tow (in seconds)
- S = ship speed during tow (in meters per second)

```{r echo = FALSE}
#constants
diameter_m <- 0.61  # diameter in meters
radius_m <- diameter_m / 2  # radius in meters
A <- pi * radius_m^2  # area of the net mouth in square meters

#ship_speed <- knots * 0.514444 # knots converted to m/s
```

```{r}
towTimes <- towTimes %>%
  filter(instrument == "Bongo Net") %>%  # Filter for instrument
  mutate(
    GEAR_VOLUME_CALCULATED = A * seconds_elapsed * (ifelse(is.na(STW_start), 
                                                          ifelse(is.na(STW_end), 1.5, abs(STW_end)), 
                                                          abs(STW_start)) * 0.514444)
  )
```

```{r echo = FALSE, include = FALSE}
towTimes_long2 <- towTimes %>%
  pivot_longer(cols = c(vol_filtered_m3_150, vol_filtered_m3_335), 
               names_to = "volume_type", 
               values_to = "vol_filtered_m3") %>%
  mutate(volume_type = recode(volume_type, 
                              vol_filtered_m3_150 = "150 µm", 
                              vol_filtered_m3_335 = "335 µm"))
towTimes_long_melted <- towTimes %>%
  pivot_longer(cols = c(vol_filtered_m3_335, 
                        vol_filtered_m3_150, 
                        GEAR_VOLUME_CALCULATED), 
               names_to = "volume_type", 
               values_to = "value") %>%
  mutate(volume_type = recode(volume_type, 
                              vol_filtered_m3_150 = "150 µm", 
                              vol_filtered_m3_335 = "335 µm"))
```


<span style="GEAR_VOLUME_CALCULATED below == volume sampled in meters squared calculated using ship speed and sampling time as specified above">red</span> text line.

```{r, warning=FALSE, echo = FALSE, fig.width=11, fig.height=6}
unique_cruises1 <- unique(towTimes_long2$cruise)
for (cruise in unique_cruises1) {
  
  # scatter plot - left panel
  scatter_plot <- ggplot(towTimes_long2 %>% filter(cruise == !!cruise & 
                                                   !is.na(vol_filtered_m3)), 
                         aes(x = GEAR_VOLUME_CALCULATED, 
                             y = vol_filtered_m3, 
                             color = volume_type, 
                             shape = volume_type)) +
    geom_abline(slope = 1, intercept = 0, 
                linetype = "dashed", color = "darkgrey", size = 1.2) +
    geom_point(size = 3.2) +
    scale_y_continuous(limits = c(0, 600)) +  
    scale_x_continuous(limits = c(0, 600)) + 
    scale_color_manual(values = c("#17356E", "#00BFB2")) +#"#755675", "#6ECC6C", "#78DF77", "#2F452E",
    labs(title = paste("Cruise:", cruise), 
         x = "Gear Volume Calculated (m³)",
         y = "Volume Filtered (m³)") +
    theme_few() +
    theme(legend.title = element_blank(),
          legend.text = element_text(size = 15, color = "black"),
          axis.title = element_text(size = 15, color = "black"), 
          axis.text = element_text(size = 14, color = "black"),
          legend.position = "bottom")
  
  # bar plot - right panel
  bar_plot <- ggplot(towTimes_long_melted %>% filter(cruise == !!cruise), 
                     aes(x = station, y = value, fill = volume_type)) +
    geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
    scale_fill_manual(values = c("#17356E", "#00BFB2", "#CD4942")) + 
    scale_y_continuous(expand = c(0, 0)) +
    labs(title = paste("Cruise:", cruise), 
         y = "Volume Filtered (m³)") +
    theme_few() +
    theme(axis.title.y = element_blank(), 
          axis.title.x = element_blank(),
          axis.text = element_text(size = 14, color = "black"),
          legend.title = element_blank(),
          legend.text = element_text(size = 15, color = "black"),
          legend.position = "bottom",
          axis.text.x = element_text(angle = 45, hjust = 1)
          )
  
  # combine plots
  combined_plot <- plot_grid(scatter_plot, bar_plot, ncol = 2)

  print(combined_plot) 
  
}
```

\newpage

## Summary

AR32 flowmeter vals look good, use those  
**AR38 L3 and L10 something off; L3 bad 335 flow; L10 bad GEAR_VOLUME_CALC**  
AR63 its okay  
AR77 L2 and L9 bad GEAR_VOLUME_CALC?; but flowmeter vals look good, use those  
AT46 flowmeter vals look good, use those  
EN608 flowmeter vals look good, use those; no times for L6-L11  
**EN617 L11 looks high but comparable to each other**  
**EN627 bad flowmeter cruise; use GEAR_VOLUME_CALC?**  
EN644 good  
EN649 good  
EN655 good  
EN657 MVCO 150 flow bad; use GEAR_VOLUME_CALC  
EN661 good  
EN668 L2, L3, L4 bad 335 flow; use GEAR_VOLUME_CALC  
**EN687 L8 the 150 flow looks suspect?, possibly use GEAR_VOLUME_CALC**   
EN695 okay  
EN706 L5 bad 150 flow; use GEAR_VOLUME_CALC  
**EN712 it seems like the 150 flow was low? at L1, L7, L11**  
**EN715 looks pretty good; 335 flow a bit high at L6**  
EN720 good  
**HRS2303 L9 flow possibly too high?**  
  
  
\newpage




### extra

```{r echo = FALSE, fig.width=12, fig.height=10}
ggplot(towTimes %>% filter(!is.na(vol_filtered_m3_335) & 
                             vol_filtered_m3_335 > 0 & 
                             station != "u11c"), 
       aes(x = station, y = vol_filtered_m3_335)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ cruise, scales = "free") +
  theme_minimal() +
  labs(title = "Free y-axis",
       y = "Volume Filtered 335 µm") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18))
ggplot(towTimes %>% filter(!is.na(vol_filtered_m3_335) & 
                             vol_filtered_m3_335 > 0 & 
                             station != "u11c"), 
       aes(x = station, y = vol_filtered_m3_335)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ cruise) +
  theme_minimal() +
  labs(title = "Fixed y-axis",
       y = "Volume Filtered 335 µm") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18))
```

```{r echo = FALSE, fig.width=12, fig.height=10}
ggplot(towTimes %>% filter(!is.na(vol_filtered_m3_150) & 
                             vol_filtered_m3_150 > 0 & 
                             station != "u11c"), 
       aes(x = station, y = vol_filtered_m3_150)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ cruise, scales = "free") +
  theme_minimal() +
  labs(title = "Free y-axis",
       y = "Volume Filtered 150 µm") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18))
ggplot(towTimes %>% filter(!is.na(vol_filtered_m3_150) & 
                             vol_filtered_m3_150 > 0 & 
                             station != "u11c"), 
       aes(x = station, y = vol_filtered_m3_150)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ cruise) +
  theme_minimal() +
  labs(title = "Fixed y-axis",
       y = "Volume Filtered 150 µm") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16),
        plot.title = element_text(size = 18))
```

